The \dealii{} Library, Version 9.3

### 2021-02-18 9.3/paper.tex

\hypersetup{
  pdfauthor={
     Daniel Arndt,
     Wolfgang Bangerth,
     Bruno Blais,
     Thomas C. Clevenger,
    Marc Fehling,
    Alexander V. Grayver,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Matthias Maier,
    Peter Munch,
    Jean-Paul Pelteret,
    Reza Rastak,
    Ignacio Thomas,
    Bruno Turcksin,
    Zhuoran Wang,
    David Wells
  },
  pdftitle={The deal.II Library, Version 9.3, 2021},
}

@@

\author[3]{Bruno Blais}
\affil[3]{Research Unit for Industrial Flows Processes (URPEI), Department of Chemical Engineering,
          Polytechnique Montréal,
          PO Box 6079, Stn Centre-Ville, Montréal, Québec, Canada, H3C 3A7.
          {\texttt{bruno.blais@polymtl.ca}}}

 \author[4]{Thomas~C.~Clevenger}
 \affil[4]{School of Mathematical and Statistical Sciences,
   Clemson University,
   Clemson, SC, 29634, USA
   {\texttt{tcleven/heister@clemson.edu}}}
%
% \author[4]{Denis~Davydov}
% \affil[4]{Chair of Applied Mechanics,
%   Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
%   Egerlandstr.\ 5,
%   91058 Erlangen, Germany.
%   {\texttt{\{denis.davydov,jean-paul.pelteret\}@fau.de}}}
%
\author[5]{Marc~Fehling}
\affil[5]{Institute for Advanced Simulation,
  Forschungszentrum J\"{u}lich GmbH,
  52425 J\"{u}lich, Germany.
  {\texttt{m.fehling@fz-juelich.de}}}
%
% \author[6]{Daniel Garcia-Sanchez}
% \affil[6]{Sorbonne Universit\'es, UPMC Univ.\ Paris 06, CNRS-UMR 7588,
%   Institut des NanoSciences de Paris, F-75005, Paris, France
%   {\texttt{daniel.garcia-sanchez@insp.upmc.fr}}}
%
% \author[2]{Graham Harper}
%

\author[6]{Alexander~V.~Grayver}
\affil[6]{Institute of Geophysics,
  ETH Zurich,
  Sonneggstrasse 5, 8092 Z\"{u}rich, Switzerland.
  {\texttt{agrayver@ethz.ch}}}

@@

 \author[12]{Reza Rastak}
 \affil[12]{Department of Civil and Environmental Engineering,
   Stanford University,
   Stanford, CA 94305, USA.
   {\texttt{rastak@stanford.edu}}}

 \author[**13]{Ignacio Tomas}
 \affil[13]{Sandia National Laboratories, Org 1442,
 P.O. Box 5800, MS 1320, Albuquerque, NM, 87185-1320,
   \texttt{itomas@sandia.gov}}

@@

\author[14]{Zhuoran Wang}
\affil[14]{Department of Mathematics, Colorado State University, Fort Collins,
          CO 80523-1874, USA.
{\texttt{zhrwang@math.colostate.edu}}}

@@

\begin{abstract}
  This paper provides an overview of the new features of the finite element
  library \dealii, version 9.2.
\end{abstract}

@@

\dealii{} version 9.2.0 was released May 20, 2020.
This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii{} software library version 9.2. \dealii{} is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL). Downloads are available at
\url{https://www.dealii.org/} and \url{https://github.com/dealii/dealii}.

The major changes of this release are:
%
\begin{itemize}
  \item A new, parallel, fully distributed triangulation class (see
        Section~\ref{subsec:pft});
  \item Substantially improved performance for very large computations
    on tens or hundreds of thousands of processor cores and up to
    trillions of unknowns (see
        Section~\ref{subsec:performance});
  \item Better support for parallel $hp$-adaptive algorithms (see
        Section~\ref{subsec:hp});
  \item Support for particle-based methods (see
        Section~\ref{subsec:particles});
  \item Improved performance of the symbolic differentiation framework (see
        Section~\ref{subsec:symbdiff});
  \item Advances in SIMD capabilities and the matrix-free infrastructure (see
        Section~\ref{subsec:mf});
  \item Advances in GPU support (see
        Section~\ref{subsec:gpu});
  \item Better use of modern C++ language features (see
        Section~\ref{subsec:cxx});
  \item Seven new tutorial programs (see
        Section~\ref{subsec:steps}).
\end{itemize}
%
These major changes are discussed in detail in Section~\ref{sec:major}. There
are a number of other noteworthy changes in the current \dealii{} release
that we briefly outline in the remainder of this section:
%
\begin{itemize}
  \item globally unique cell IDs
\end{itemize}
%
The changelog lists more than 240 other
features and bugfixes.

@@

\newpage

@@

@@

In addition, there are a number of new tutorial programs:
\begin{itemize}
\item \texttt{step-19} TODO
\item \texttt{step-67b} is an explicit time integrator for the
        compressible Euler equations discretized with a high-order discontinuous
        Galerkin (DG) scheme, using the matrix-free infrastructure just as \texttt{step67} does.
        The tutorial presents advanced topics, like the usage of cell-centric loops and
        the new MPI-3.0 shared-memory capabilities of \texttt{MatrixFree} tor each high
        throughput. Furthermore, the utilization of the template parameter
        \texttt{VectorizedArrayType} and the application of lambdas to capture cell and face
        integrals are discussed.
\item \texttt{step-68} TODO
\item \texttt{step-75} TODO
\end{itemize}

There are also new programs in the code gallery (a collection of
user-contributed programs that often solve more complicated problems
than tutorial programs, and intended as starting points for further
research rather than as teaching tools):
\begin{itemize}
  \item 
\end{itemize}

@@

\dealii{} is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people
contributed code to this release:\\
%
% Uwe Koecher doesn't usually show up in the changelog, but
% we should make sure he's listed.
%
% 9.2: updated 5/11/2020 MM
Pasquale Africa,
Ashna Aggarwal,
Giovanni Alzetta,
Mathias Anselmann,
Kirana Bergstrom,
Manaswinee Bezbaruah,
Benjamin Brands,
Yong-Yong Cai,
Fabian Castelli,
Joshua Christopher,
Ester Comellas,
Katherine Cosburn,
Denis Davydov,
Elias Dejene,
Stefano Dominici,
Brett Dong,
Luel Emishaw,
Niklas Fehn,
Isuru Fernando,
Rebecca Fildes,
Menno Fraters,
Andres Galindo,
Daniel Garcia-Sanchez,
Rene Gassmoeller,
Melanie Gerault,
Nicola Giuliani,
Brandon Gleeson,
Anne Glerum,
Krishnakumar Gopalakrishnan,
Graham Harper,
Mohammed Hassan,
Nicole Hayes,
Bang He,
Johannes Heinz,
Jiuhua Hu,
Lise-Marie Imbert-Gerard,
Manu Jayadharan,
Daniel Jodlbauer,
Marie Kajan,
Guido Kanschat,
Alexander Knieps,
Uwe K{\"o}cher,
Paras Kumar,
Konstantin Ladutenko,
Charu Lata,
Adam Lee,
Wenyu Lei,
Katrin Mang,
Mae Markowski,
Franco Milicchio,
Adriana Morales Miranda,
Bob Myhill,
Emily Novak,
Omotayo Omosebi,
Alexey Ozeritskiy,
Rebecca Pereira,
Geneva Porter,
Laura Prieto Saavedra,
Roland Richter,
Jonathan Robey,
Irabiel Romero,
Matthew Russell,
Tonatiuh Sanchez-Vizuet,
Natasha S. Sharma,
Doug Shi-Dong,
Konrad Simon,
Stephanie Sparks,
Sebastian Stark,
Simon Sticko,
Jan Philipp Thiele,
Jihuan Tian,
Sara Tro,
Ferdinand Vanmaele,
Michal Wichrowski,
Julius Witte,
Winnifried Wollner,
Ming Yang,
Mario Zepeda Aguilar,
Wenjuan Zhang,
Victor Zheng.

>>>>

\hypersetup{
  pdfauthor={
     Daniel Arndt,
     Wolfgang Bangerth,
     %Bruno Blais,
     %Thomas C. Clevenger,
    Marc Fehling,
    %Alexander V. Grayver,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Matthias Maier,
    Peter Munch,
    Jean-Paul Pelteret,
    %Reza Rastak,
    %Ignacio Thomas,
    Bruno Turcksin,
    %Zhuoran Wang,
    David Wells
  },
  pdftitle={The deal.II Library, Version 9.3, 2021},
}

@@

%\author[3]{Bruno Blais}
%\affil[3]{Research Unit for Industrial Flows Processes (URPEI), Department of Chemical Engineering,
%          Polytechnique Montréal,
%          PO Box 6079, Stn Centre-Ville, Montréal, Québec, Canada, H3C 3A7.
%          {\texttt{bruno.blais@polymtl.ca}}}

% \author[4]{Thomas~C.~Clevenger}
% \affil[4]{School of Mathematical and Statistical Sciences,
%   Clemson University,
%   Clemson, SC, 29634, USA
%   {\texttt{tcleven/heister@clemson.edu}}}
%
% \author[4]{Denis~Davydov}
% \affil[4]{Chair of Applied Mechanics,
%   Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
%   Egerlandstr.\ 5,
%   91058 Erlangen, Germany.
%   {\texttt{\{denis.davydov,jean-paul.pelteret\}@fau.de}}}
%
\author[5]{Marc~Fehling}
\affil[5]{Institute for Advanced Simulation,
  Forschungszentrum J\"{u}lich GmbH,
  52425 J\"{u}lich, Germany.
  {\texttt{m.fehling@fz-juelich.de}}}
%
% \author[6]{Daniel Garcia-Sanchez}
% \affil[6]{Sorbonne Universit\'es, UPMC Univ.\ Paris 06, CNRS-UMR 7588,
%   Institut des NanoSciences de Paris, F-75005, Paris, France
%   {\texttt{daniel.garcia-sanchez@insp.upmc.fr}}}
%
% \author[2]{Graham Harper}
%

%\author[6]{Alexander~V.~Grayver}
%\affil[6]{Institute of Geophysics,
%  ETH Zurich,
%  Sonneggstrasse 5, 8092 Z\"{u}rich, Switzerland.
%  {\texttt{agrayver@ethz.ch}}}

@@

% \author[12]{Reza Rastak}
% \affil[12]{Department of Civil and Environmental Engineering,
%   Stanford University,
%   Stanford, CA 94305, USA.
%   {\texttt{rastak@stanford.edu}}}

% \author[**13]{Ignacio Tomas}
% \affil[13]{Sandia National Laboratories, Org 1442,
% P.O. Box 5800, MS 1320, Albuquerque, NM, 87185-1320,
%   \texttt{itomas@sandia.gov}}

@@

%\author[14]{Zhuoran Wang}
%\affil[14]{Department of Mathematics, Colorado State University, Fort Collins,
%          CO 80523-1874, USA.
%{\texttt{zhrwang@math.colostate.edu}}}

@@

\begin{abstract}
  This paper provides an overview of the new features of the finite element
  library \dealii, version 9.3.
\end{abstract}

@@

\dealii{} version 9.3.0 was released May ??, 2021.
This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii{} software library version 9.3. \dealii{} is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL). Downloads are available at
\url{https://www.dealii.org/} and \url{https://github.com/dealii/dealii}.

The major changes of this release are:
%
\begin{itemize}
  \item TODO.
\end{itemize}
%
These major changes are discussed in detail in Section~\ref{sec:major}. There
are a number of other noteworthy changes in the current \dealii{} release
that we briefly outline in the remainder of this section:
%
\begin{itemize}
  \item globally unique cell IDs: TODO
\end{itemize}
%
The changelog lists more than 240 other
features and bugfixes.

@@

%\newpage

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Advances in the particle infrastructure}
\label{subsec:particles}

TODO

@@

In addition, there are a number of new tutorial programs:
\begin{itemize}
\item \texttt{step-19} TODO
\item \texttt{step-68} TODO
\item \texttt{step-75} TODO
\item \texttt{step-76} is an explicit time integrator for the
        compressible Euler equations discretized with a high-order discontinuous
        Galerkin (DG) scheme, using the matrix-free infrastructure just as \texttt{step67} does.
        The tutorial presents advanced topics, like the usage of cell-centric loops and
        the new MPI-3.0 shared-memory capabilities of \texttt{MatrixFree} tor each high
        throughput. Furthermore, the utilization of the template parameter
        \texttt{VectorizedArrayType} and the application of lambdas to capture cell and face
        integrals are discussed.
\end{itemize}

There are also new programs in the code gallery (a collection of
user-contributed programs that often solve more complicated problems
than tutorial programs, and intended as starting points for further
research rather than as teaching tools):
\begin{itemize}
  \item preCICE: TODO
\end{itemize}

@@

\dealii{} is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people
contributed code to this release:\\
%
% Uwe Koecher doesn't usually show up in the changelog, but
% we should make sure he's listed.
%
% 9.2: updated 5/11/2020 MM
TODO
